---
title: "HW 8"
author: "Anish Palley, ap65675, https://github.com/anishpalley/SDS-315-HW8"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
  pdf_document: default
---

```{r global options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, # show the code
                      eval = TRUE, # run the code
                      warning = FALSE, # show warning messages
                      message = FALSE, # show basic messages
                      fig.align = "center"
                      )
#Uploading necessary packages
library(tidyverse)
library(ggplot2)
library(rvest)
library(dplyr)
library(plotly)
library(mosaic)
library(MatchIt)
```

## Problem 1: Armfolding

```{r}
creatinine <- read_csv("creatinine.csv")
```

### Part A)
```{r}
model_cr = lm(creatclear ~ age, data=creatinine)
clearance_rate_55 = -0.6198159*55 + 147.8129158
clearance_rate_55
```
We would expect a 55-year-old to have a creatinine clearance rate of 113.723 mL/minute by using a linear regression analysis. With a rounded y-intercept of 147.813 and a rounded slope of -0.620, we can multiple the slope by the given age, 55, and add it to the y-intercept resulting in a creatinine clearance rate of 113.723 mL/minute.

### Part B)
```{r}
coef(model_cr)
```
Using the slope of the linear regress line, for every increase in age by a year, the creatinine clearance rate decreases by -0.620 mL/minute per year.

### Part C)
```{r}
actual_40 = 135
actual_60 = 112
predicted_40 = -0.6198159*40 + 147.8129158
predicted_60 = -0.6198159*60 + 147.8129158
residual_40 = actual_40-predicted_40
residual_60 = actual_60-predicted_60
residual_40
residual_60
```
To adjust for the systematic effect of age, we use fair comparison to find the creatinine clearance rate residuals for the 40-year-old and the 60-year-old. The 40-year-old creatinine clearance rate residual is 11.98 compared to the 60-year-old creatinine clearance rate residual of 1.38. The 40-year-old with a rate of 135 mL/minute has a higher (more positive) residual, indicating that they have a healthier creatinine clearance rate for their age.

## Problem 2: Modeling disease growth
```{r}
covid <- read_csv("covid.csv")
```
### 1.
```{r}
italy <- covid |>
  filter(country == "Italy")
#model_italy <- lm(log(deaths) ~ #log(days_since_first_death), data = italy)
boot_italy = do(3)*lm(log(deaths) ~ (days_since_first_death), data = resample(italy))
confint(boot_italy, level = .95)
growth_rate <- 0.183218
doubling_time <- 70/(growth_rate*100)
round(growth_rate,3)
round(doubling_time, 1)
```
The estimated growth rate of the daily death total for Italy is 0.183, with a doubling time of 3.8 days. We are 95% confident that the true growth rate of the daily death total for Italy is somewhere between 0.159 and 0.208, thus the true doubling rate somewhere between 4.402 and 3.365 days.
```{r}
model_italy = lm(log(deaths) ~ (days_since_first_death), data = italy)
coef(model_italy)
```

### 2.
```{r}
spain <- covid |>
  filter(country == "Spain")
#model_italy <- lm(log(deaths) ~ #log(days_since_first_death), data = italy)
boot_spain = do(10000)*lm(log(deaths) ~ (days_since_first_death), data = resample(spain))
confint(boot_spain, level = .95)
growth_rate <- 0.2762447
doubling_time <- 70/(growth_rate*100)
round(growth_rate,3)
round(doubling_time, 1)
```
The estimated growth rate of the daily death total for Spain is 0.276, with a doubling time of 2.5 days. We are 95% confident that the true growth rate of the daily death total for Spain is somewhere between 0.235 and 0.318, thus the true doubling rate somewhere between 2.979 and 2.201 days.

```{r}
lm_spain = lm(log(deaths) ~ days_since_first_death, data=spain)
coef(lm_spain)
```
### 3.
```{r}
ggplot(covid) + geom_line(aes(x=days_since_first_death, y=deaths, color = country))+
  labs(
    x = "Days since first death",
    y = "Deaths",
    title = "Daily deaths over time in Italy and Spain"
  )
```
## Problem 3: price elasticity of demand
```{r}
milk <- read_csv("milk.csv")
```

```{r}
boot_milk <- do(10000) * lm(log(sales) ~ log(price), data = resample(milk))
confint(boot_milk, level = .95)
```
The estimated price elasticity of demand for milk is rounded to -1.62. We are 95% confident that the true price elasticity of demand for milk is somewhere between -1.769 and -1.460. After logging both sales and price in order to obtain a linear relationship, we bootstrapped 10000 samples of a linear regression between log(milk sales) and log(milk prices). To interpret the price elasticity of demand value, we say that a 1% increase in milk prices results in consumers wanting to buy about 1.62% less of it, on average.
```{r}
lm_milk = lm(log(sales) ~ log(price), data=milk)
coef(lm_milk)
```